---
title: "GDD"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GDD}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
pacman::p_load("DT", "tidyverse", "stotzgdh")
```


# Prior Knowledge Needed

* Ability to understand and load files in R
* Basic Algebra

# What is GDD and GDH

* GDD (Growing Degree Days) is the measurement of heat units that a plant has absorbed throughout a day. 
* GDH (Growing Degree Hours) is the same measurement as GDD except it measures it by hour than by day.

## Why calculate GDD/GDH

* Measuring GDD or GDH can help track the developmental cycle of most crops. 

## How is GDD/GDH calulated

* For each plant, there is a threshold temperature by which a plant needs in order to start absorbing heat.
* Plants also have a maximum temperature at which the plant cannot absorb anymore heat.
* The steps are listed below:
  1. For GDH, take the temperature listed at the hour
  2. Compare the temperature to the plant's high and low thresholds
    - If the temperature is below the high and above low, keep the temperature
    - If the temperature is above the high threshold, keep the high threshold temperature
    - If the temperature is below the low threshold, keep the low threshold temperature
  3. Take the keep temperature and subtract it by the low threshold temperature
  4. For GDD, sum all the GDH calculations and then divide by the low temperature threshold.
* There are other methods for calculating GDD, but the current method works best for the data pulled from the data base.

## Formulas 

$$
\text{GDH}_i = \cases{ \text{T}_i - \text{BT}   & $\text{when HT} > \text{T}_i > \text{BT}$ \cr
                      \text{HT} - \text{BT}     & $\text{when T}_i>\text{HT}$ \cr
                      0                         & $\text{when T}_i < \text{BT}$}\\
\text{Where } \\
\text{T is the temperature,} \\ 
\text{BT is the base temperature thresholdof the crop,} \\
\text{HT is the high temperature threshold of the crop,}
\text{and } i \text{ is the hour the temperature is taken.}  
$$


$$
\text{GDD} = \frac{\sum{\text{GDH}}}{\text{Base Temperature}} = \frac{\text{GDH}_1 + \text{GDH}_2 + ... + \text{GDH}_{24}}{\text{Base Temperature}}
$$

# The Fucnction `gdh()`

* The `gdh()` function calculates the GDH for each of the hourly temperatures found within the temperature file and then returns the GDD for each of the days given in the original file.

## What is required to run the GDH function in `stotzgdh` 

* temperature by hours - this is taken from the HRRR data
* crop threshold temperature file - this is taken from file `crop_base_temp.rda` found in the package
* crop name - this is taken from both the file imputed by the user earlier and `crop_base_temp.rda`
  - if the crop names in the imputed file does not match any of the crops found in `crop_base_temp.rda`, the package  will prompt the user to match the crop names not found in `crop_base_temp.rda` by showing the crop name that is not found and examples of crops that can be used and asks the user to correct their imputed file. 

## Example

For this example, I will be only running one row of a fake dataset that I created to show how the function works. I first load in the data and display the table of fake data below. 

```{r}
field_id <- c('2c1aed90-3cb6-4476-b237-44162245c8cb')
crop_season <- c(2022)
crop_name <- c("cotton")
lat <- c(33.049406331)
long <- c(-112.078917325)
seeding_date <- c("4/13/2022")
harvest_date <- c("10/27/2022")
hour0 <- c(56)
hour1 <- c(54)
hour2 <- c(53)
hour3 <- c(51)
hour4 <- c(49)
hour5 <- c(48)
hour6 <- c(47)
hour7 <- c(47)
hour8 <- c(52)
hour9 <- c(56)
hour10 <- c(59)
hour11 <- c(62)
hour12 <- c(65)
hour13 <- c(67)
hour14 <- c(70)
hour15 <- c(73)
hour16 <- c(74)
hour17 <- c(75)
hour18 <- c(76)
hour19 <- c(73)
hour20 <- c(69)
hour21 <- c(67)
hour22 <- c(64)
hour23 <- c(60)

temp_data_hours <- data.frame(crop_name,lat,long,hour0,hour1,hour2,hour3,hour4,hour5, hour6,
                   hour7,hour8,hour9,hour10,hour11,hour12,hour13,
                   hour14,hour15,hour16,hour17,hour18,hour19,
                   hour20,hour21,hour22,hour23)
```


```{r}
# test_dat <- stotzgdh::temp_data_hours

DT::datatable(temp_data_hours)
```

Next, I will set up a new data set similar to the script for the package, adding a column for GDD. I will then call the function to re-write that `NA` value. For bigger data sets, as seen in the main script, the use of for loop is recommended. The number below shows the value of the calculated GDD from the fake data set. 

```{r}
result <- temp_data_hours
result$GDD <- NA


result$GDD[1] <- stotzgdh::gdh(result[1, 4:27], crop = result$`crop_name`[1], crop_data = stotzgdh::crop_base_temp)
   
 
print(result$GDD)
```


