---
title: "GDD"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GDD}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#library(stotzgdh) #does not load
pacman::p_load("DT", "tidyverse")
```


```{r gdh}
gdh <- function(..., crop, crop_data) {
  crop_subset <- subset(crop_data, crop_name == crop)
  # Obtain base temp value from crop_data based on crop
  base_temp <- as.numeric(crop_subset[1,2])
  # Obtain upper threshold value from crop_data based on crop
  threshold <- as.numeric(crop_subset[1,4])
  # Check for NA values using is.na() and return NA if any temperature values are missing
  if (any(is.na(c(...)))) {
    return(NA)
  } else {
    # Replace negative values with zero using pmax()
    # Replace values exceeding the threshold with the threshold value
    values <- pmax(pmin(..., threshold)-base_temp, 0)
    
 
    # Return the sum of the values and divide by the base temp
    return(sum(values)/base_temp)
  }
}
```

# Prior Knowledge Needed

* Ability to understand and load files in R
* Basic Algebra

# What is GDD and GDH

* GDD (Growing Degree Days) is the measurement of heat units that a plant has absorbed throughout a day. 
* GDH (Growing Degree Hours) is the same measurement as GDD except it measures it by hour than by day.

## Why calculate GDD/GDH

* Measuring GDD or GDH can help track the developmental cycle of most crops. 

## How is GDD/GDH calulated

* For each plant, there is a threshold temperature by which a plant needs in order to start absorbing heat.
* Plants also have a maximum temperature at which the plant cannot absorb anymore heat.
* The steps are listed below:
  1. For GDH, take the temperature listed at the hour
  2. Compare the temperature to the plant's high and low thresholds
    - If the temperature is below the high and above low, keep the temperature
    - If the temperature is above the high threshold, keep the high threshold temperature
    - If the temperature is below the low threshold, keep the low threshold temperature
  3. Take the keep temperature and subtract it by the low threshold temperature
  4. For GDD, sum all the GDH calculations and then divide by the low temperature threshold.
* There are other methods for calculating GDD, but the current method works best for the data pulled from the data base.

## Formulas 

$$
\text{GDH}_i = \text{Temperature}_i - \text{Base Temperature}\\
\text{Where } i \text{ is the hour the temperature is taken.}  
$$


$$
\text{GDD} = \frac{\sum{\text{GDH}}}{\text{Base Temperature}} = \frac{\text{GDH}_1 + \text{GDH}_2 + ... + \text{GDH}_{24}}{\text{Base Temperature}}
$$

# The Fucnction `gdh()`

* The `gdh()` function calculates the GDH for each of the hourly temperatures found within the temperature file and then returns the GDD for each of the days given in the original file.

## What is required to run the GDH function in `stotzgdh` 

* temperature by hours - this is taken from the HRRR data
* crop threshold temperature file - this is taken from file `crop_base_temp.rda` found in the package
* crop name - this is taken from both the file imputed by the user earlier and `crop_base_temp.rda`
  - if the crop names in the imputed file does not match any of the crops found in `crop_base_temp.rda`, the package  will prompt the user to match the crop names not found in `crop_base_temp.rda` by showing the crop name that is not found and examples of crops that can be used and asks the user to correct their imputed file. 

## Example

```{r}
load("~/GitHub/package/stotzgdh/data/crop_base_temp.rda")
DT::datatable(crop_base_temp)
```

